---
title: "Data Wrangling with Tidyverse"
author: "Niklas Felix Cypris"
output: pdf_document
---

```{r setup}
library(tidyverse)
library(visdat)

Sys.setenv(LANG = "en")
```

# TOPICS

1. FILTERING ROWS

2. SELECTING COLUMNS

3. CREATING NEW VARIABLES

4. SUMMARIZING COLUMNS

5. GROUPING OBSERVATIONS

6. JOINING DATA SETS

# READ DATA

```{r}
df <- read_csv(file = "./data/movies.csv")

bechdel <- read_csv(file = "./data/bechdel.csv")
```

```{r}

```

# FILTERING ROWS

filter() lets you pick rows based on the conditions you input.

## Single condition

x < y - x smaller than y

x > y - x bigger than y

x >= y - x bigger than or equal to y

x == y - x equal to y

x != y - x unequal to y

x %in% c(a, b, c) - x equal to  a, b, or c

is.na(x) - x is NA

!is.na(x) - x is not NA


Get only movies with production budget above $100,000,000

```{r}

```

Creating a subset

```{r}

```

```{r}

```

## Multiple conditions

X & Y - both conditions must be fulfilled

X | Y - one of the conditions must be fulfilled

```{r}

```

```{r}

```

## Exercise

Keep only movies that were released after 2010. How many observations do you have?

```{r}

```

Keep only rows where the PG rating is "PG-13" and which belong to the "Comedy", "Horror", or "Drama" genre.

```{r}

```

# SELECTING COLUMNS

select() lets you pick columns based on the conditions you input.

Get only the release_date and genre columns

```{r}

```

Get everything but the release_date and genre columns

```{r}

```

## Useful helper functions

starts_with():

```{r}

```

ends_with():

```{r}

```

contains():

```{r}

```

everything():

```{r}

```

## Exercise

Move the genre column to the front.

```{r}

```

Select all columns in the bechdel data set that measure movie budget.

```{r}

```

Remove all columns from the bechdel data set that indicate values adjusted to 2013 prices.

```{r}

```

# CREATING NEW VARIABLES

Creating new variables

```{r}

```

```{r}

```

## sum() & rowSums() | mean() & rowMeans()

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```

## Exercise

Create a column for the bechdel data set that computes how old the movies are. Move that column to the front of the data set.

```{r}

```

Compute the overall boxoffice success adjusted to 2013 prices (made up of domgross_2013 and intgross_2013) for each movie in the bechdel data set.

```{r}

```

Create a column called "pass" that says TRUE when the "clean_test" variable for an observation is "ok" and FALSE when the variable is anything else. Think of the logical operators that were introduced above.

```{r}

```

# SUMMARIZING COLUMNS

Looking at summary values of columns

```{r}

```

```{r}

```

# GROUPING OBSERVATIONS

Looking at subgroups in the data

e.g. average international box office success per year

```{r}

```

## Exercise

Filter only movies between 2010 and 2013 and look at the average domestic gross per year and genre (input both variables in the group_by-function at the same time).

```{r}

```

Compute the overall returns of movies per year starting 1990

```{r}

```

# JOINING DATA SETS

## CBIND() & RBIND()

```{r}
bechdel_col1 <- bechdel[,1:5]
bechdel_col2 <- bechdel[,6:ncol(bechdel)]
```

```{r}

```

```{r}

```

```{r}
bechdel_row1 <- bechdel[1:1000,]
bechdel_row2 <- bechdel[1001:nrow(bechdel),]
```

```{r}

```

```{r}

```

## JOINS

https://github.com/gadenbuie/tidyexplain

```{r make example data sets}
bechdel <- bechdel %>% 
  mutate(id = 1:nrow(.)) %>% 
  select(id, everything())

bechdel1 <- bechdel[1:(nrow(bechdel)-4),1:4]
bechdel2 <- bechdel[5:nrow(bechdel),c(1, 5:ncol(bechdel))]
```

## INNER JOIN

All rows from x where there are matching values in y, and all columns from x and y.

```{r}

```

## LEFT JOIN

All rows from x, and all columns from x and y. Rows in x with no match in y will have NA values in the new columns.

```{r}

```

## RIGHT JOIN

All rows from y, and all columns from x and y. Rows in y with no match in x will have NA values in the new columns.

```{r}

```

## FULL JOIN

All rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing.

```{r}

```

## SEMI JOIN

All rows from x where there are matching values in y, keeping just columns from x.

```{r}

```

```{r}

```

## ANTI JOIN

All rows from x where there are not matching values in y, keeping just columns from x.

```{r}

```

## EXERCISE

```{r make example data sets}
movies <- df %>% 
  mutate(id = 1:nrow(.)) %>% 
  select(id, everything())

movies1 <- movies[1:(nrow(movies)-400), 1:3]
movies2 <- movies[200:nrow(movies), c(1, 4:ncol(movies))]
```

You want to join movies1 and movies2 to create a movies_new1 data frame. That data frame should contain all rows and columns from movies1 and all columns from movies2. Select the correct *join() command.

```{r}

```

You want to join movies1 and movies2 to create a movies_new2 data frame. That data frame should contain all rows and columns from movies1 and all rows and columns from movies2. Select the correct *join() command.

```{r}

```