---
title: "read_data_examples"
author: "Niklas Cypris"
output: pdf_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

# Bechdel Data Set

https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-09

# IMPORTING AND EXPORTING DATA

## Csv

### Comma-separated variables

Read

```{r}
df_csv1 <- read_csv(file = "./data/bechdel_csv1.csv")
```

Write

```{r}
write_csv(df_csv1, file = "./data/bechdel_csv1.csv")
```

### Semi-colon separated variables

Decimal separators vary between countries. When commas serve as decimal separators, semi-colons are used as variable separators in csv files.

https://en.wikipedia.org/wiki/Decimal_separator#/media/File:DecimalSeparator.svg

Read

```{r}
df_csv2 <- read_csv2(file = "./data/bechdel_csv2.csv")
```

Write

```{r}
write_csv2(df_csv2, file = "./data/bechdel_csv2.csv")
```

further reading: https://r4ds.had.co.nz/data-import.html

documentation: https://readr.tidyverse.org/


## Excel

Use `readxl` package to read excel files.

```{r}
df_xlsx <- read_xlsx(path = "./data/bechdel_xlsx.xlsx")
```

Read first sheet. Specify columns and rows

documentation: https://readxl.tidyverse.org/

## SPSS (sav)

Use `haven` package to read sav files.

Read

```{r}
df_sav <- read_sav(file = "./data/bechdel_sav.sav")
```

Write

```{r}
write_sav(df_sav, path = "./data/bechdel_sav.sav")
```

Haven also reads Stata and Sas files.

documentation: https://haven.tidyverse.org/

## Rda files

```{r}
load("./data/bechdel_rda.Rda")
```

```{r}
save(df_rda, file = "./data/bechdel_rda.Rda")
```
